---
title: "Capstone project kmeans"
author: "Shivani Gupta"
date: "01/05/2021"
output: word_document
---

####################################### What you will learn and Implement #######################################

# > Understand kmeans clustering.
# > Understand retail Data for applying kmeans clustering.
# > Understand the objective.
# > Maths of kmeans clustering.
# > understand Optimum number of cluster.
# > Using elbow method to determine the optimal number of cluster.
# > model fitting
# > Evaluation

# > Objective

# The client is a leading Fashion retailer in Australia. Company runs a display advertising campaign for this brand, where it shows ads to users leading them to make a purchase on the brands website.The given dataset is the Sales data for all users who made a purchase online in the first half of October '17.

# Based on your analysis, arrive at a statistical segmentation of the brands audience based 
# on Revenue. The number of segments is up to you. Please provide definitions of each groups

# > desired Library installation

```{r}
# install.packages("factoextra")
# install.packages("rpivotTable")
# install.packages("cluster")
# install.packages("fpc")
# install.packages("factoextra")
# install.packages("rpivotTable")
# install.packages("ggplot2")
# install.packages("dplyr")
# install.packages("magrittr")
```


```{r}
library(ggplot2)
library(dplyr)
library(magrittr)
library(cluster)
library(fpc)
library(factoextra)
library(rpivotTable)
```

# > Setting working directory

```{r}
path<-"D:/data science/Industry Application - Kmeans"
setwd(path)
```

# > reading data

```{r}
Revenue<-read.csv("RevenueCluster.csv", header=T, na.strings=c("","NA"))
head(Revenue, 10)
str(Revenue)
Revenue$timestamp<-as.POSIXct(Revenue$timestamp, format = "%d-%m-%Y %H:%M")
```

# > subsetting/manipulating data to fit kmeans clustering

# Calculating Missing value 

```{r}
sum(is.na(Revenue))
sapply(Revenue, function(x) {
  sum(is.na(x))
})
```

#Treating missing categorical values with their mode.

```{r}
Mode<-function(v){
   uniqv <- unique(Revenue$City)
   uniqv[which.max(tabulate(match(Revenue$City, uniqv)))]
}
result<-Mode(Revenue$City)
print(result)
# mode value for city is Sydney.
# Replacing NA's with Sydney
Revenue$City[is.na(Revenue$City)]<-"Sydney"
```


```{r}
Mode2<-function(v){
   uniqv2 <- unique(Revenue$Country_Province)
   uniqv2[which.max(tabulate(match(Revenue$Country_Province, uniqv2)))]
}
result2<-Mode2(Revenue$Country_Province)
print(result2)

#mode value for Country Province is NSW
# Replacing NA's with NSW
Revenue$Country_Province[is.na(Revenue$Country_Province)]<-"NSW"
```

#Checking the data and data types
```{r}
str(Revenue)
summary(Revenue)
```

# > fitting basic model taking a random number of cluster say "3"

```{r}
set.seed(12345)
model_km<-kmeans(Revenue$Revenue.AUD., centers = 3) # Elbow methods
aggregate(Revenue$Revenue.AUD.,by=list(model_km$cluster),FUN=mean)
```
*Cluster1- Customers with Low Value*
*Cluster2- Customers with High Value*
*Cluster3- Customers with Moderate Value*

# Customers with High Value are regular in nature and generate high revenue
# Customers with Moderate Value are potential customers
# Customers with low Value are irregular or gone away in nature



# > calculating optimum number cluster

```{r}
optimcluster <- function(k){
  cluster <- kmeans(Revenue$Revenue.AUD., k)
  return(cluster$tot.withinss)
}

wss <- sapply(1:20, optimcluster)

optimK <- data.frame(k = 1:20, wss)
head(optimK, 10)
```

Interpretation: The method consists of plotting the explained variation as a function of the number of clusters, and picking the elbow of the curve as the number of clusters to use.

### ploting the k value from 1 to 20 against within sum of square (wss)

```{r}
ggplot(data = optimK) + aes(x=k, y= wss) + geom_point() + geom_line()+
  scale_x_continuous(breaks = seq(1,20,by = 1)) +geom_vline(xintercept = 4, linetype = 2)+
  labs(subtitle = "Elbow method")

OptimumClusters<- "Optimum number of cluster is 4"
```

# > Model fitting using optimum k

```{r}
set.seed(500000)
model_km_optm<-kmeans(Revenue$Revenue.AUD., centers = 4) # Elbow methods
plotcluster(Revenue$Revenue.AUD., model_km_optm$cluster)


model_km_optm$centers #Display cluster centers
Cluster_size<-table(model_km_optm$cluster) #Give a count of data points in each cluster
print(Cluster_size)
names(Cluster_size)<-c("High Value","Deal seekers","Loyal","Infrequent")
Cluster_size
Aggregate<-aggregate(Revenue$Revenue.AUD.,by=list(model_km_optm$cluster),FUN=mean)
Aggregate$Group.1<-plyr::mapvalues( Aggregate$Group.1,
                                      from = c(1,2,3,4),
                                      to = c("High Value", "Deal seekers","Loyal","Infrequent"))

Revenue$Clusters <- model_km_optm$cluster
Revenue$Groups <- plyr::mapvalues( Revenue$Clusters,
                                      from = c(1,2,3,4),
                                      to = c("High Value", "Deal seekers","Loyal","Infrequent"))
head(Revenue, 20)
```
**Analysis**

# High value customers are those customers who generates high values of revenue for the retailer in range of 892 to 5387 AUD
# Deal Seekers are customers who needs a little push and incentives in terms of sales and deals. They generate revenue in range of 135 to 325 AUD
# Loyal customers are the potential customers who are regular in purchasing and generate revenue in range of 326 to 890 AUD
# Infrequent customers are those customers who have either not generated any income or have brought revenue to the retailer less than 138 AUD

```{r}
o=order(model_km_optm$cluster)
Revenue_1<-Revenue$Revenue.AUD.[o]
Cluster_1<-model_km_optm$cluster[o]
Groups_1<- Revenue$Groups[o]
Customer_segmentation<-data.frame(Revenue_1, Cluster_1, Groups_1)
```


```{r}
ggplot(Customer_segmentation, aes(x=Groups_1, y= Revenue_1, col = as.factor(Cluster_1))) +
  geom_point()
```


```{r}
library(RColorBrewer)
coul2 <- brewer.pal(4, "Spectral")
ggplot(Revenue, aes(x=Groups))+geom_bar(fill= coul2 )+xlab("Customer Groups")+ylab("Frequency")+
ggtitle("Customer Target Groups")+ geom_text(aes(label = ..count..), stat = "count", vjust = 1.5, colour = "black")
```


```{r}
Aggregate$x<-as.numeric(Aggregate$x)
str(Aggregate$x)
coul1 <- brewer.pal(4, "Set2")
barp<-barplot(Aggregate$x, names.arg =Aggregate$Group.1, col=coul1, xlab = "Groups", ylab= "Average Revenue")

```
```{r}
write.csv(Customer_segmentation, 'Customer_segmentation.csv')

```
